\section{Motivation} \label{sec:motivation}

%%%
%
% Outline:
% - Current standard: terrestrial networking technology
%	- [Add papers on wired networking technology - glass fiber]
%	- "gold standard": Glass fiber (transmits data at the speed of light)
%	- problem: expensive infrastructure, country borders, prone to risk [Paper] --> easy to attack wiring
%	- idea: solve most of these problems by putting up satellites which take over communication
%	- depending on height of satellite, it covers a big part of the earth's population
%	- with an antenna only, one could receive and transmit messages easily
% - Good idea, and governments already put up their networked satellite systems in order to have a fallback
%   in case some disaster appears.
%	- but not intended for the broad population
%	- intended for emergencies where little people use the system
% - in recent times, also businesses tried to use this idea and put their system up in the space, e.g,. HughesNet, ViaSat and Starlink
%	- but little is known, how well that actually works
%	- it was found that the height of satellites is the first influence
%		- GEO Satellites: Satellites at very high altitude
%		- LEO Satellites: Satellites at very low altitude
%	- Usually, GEO Satellites have a very high latency and cannot compete with LEO systems
%		- [Papers with Numbers for GEO and LEO]
%		- [Add papers that measure Starlink]
%	- However, little research is available to state clearly the performance and resilience of networked satellite systems
%		- therefore, the thesis proposed here wants to determine the current state of several networked satellite systems
%		- puts special remark on Starlink, as RIPE has many probes available for that in their measurement network
%
%%%

The internet plays a major role in our everyday's life. There are many technologies available that allow to connect to the web addressing different use cases (e.g., Wi-Fi for local area networks \cite{Henry2002} or \ac{FTTH} for high-speed wired connections to an end user's home \cite{Aleksic2010}).
However, there are several problems with such a setup. First, it requires the wired infrastructure, which is quite expensive. For single end users, it is not affordable. Second, should the wired infrastructure be in place, it is vulnerable to attacks or natural catastrophes that render the service unusable.

Another emerging technology uses satellites as data transmission medium. It is very resilient to human attacks or natural catastrophes as it is largely inaccessible. Therefore, governments installed dedicated satellite constellations to maintain communication in any given scenario. A satellite constellation is a group of artificial satellites that serve a specific purpose. Prominent examples of governmental satellite constellations are
