\section{Usecase of Networked Satellite Systems} \ref{sec:usecase-networked-satellite-systems}

Networked sallite systems are around for quite a while since humanity posses both satellites and the internet.
The internet has become a key-technology for communication in nearly every aspect of the society.
Having no access to the internet will result in significant drawbacks. 
However, internet requires a complex infrastructure with high cost. Especially in distant location with little population, building such an infrastructure will not be affordable.

% TODO: Add citation for the statement "only three satellites are required"
Therefore, people came up with the idea of using satellites to communicate with distant locations.
In theory, only three satellites are required to communicate with any point on earth. 
The cost of providing this number of satellites is much less compared to the costs of providing a terrestrial network infrastructure with similar accessibility.

There is one group of people that will not receive terrestrial network infrastructure: people on boats and planes.
While there is cellular internet access, it is not offered on the sea and in the high sky.
Therefore, the only option of internet access is satellite internet, which is served all over the world.

Additionally, networked satellite systems are much more resilient to physical influences like earthquakes, terrorist attacks, or storms \cite{DBLP:conf/pam/StevensIBD24}.
Terrestrial infrastructure can be destroyed easily and therefore especially governments contracted networked satellite system ISPs.


