\subsection{Influence of Solarmagnetic Storms on Latency}
\label{sec:solarmagnetic-storms}

Ma et al. \cite{DBLP:conf/infocom/MaCZCML23} claimed solar magnetic storms have
a significant impact on the performance of Starlink. To check on that, we
conducted a study to see an actual correlation. We used the TLS handshake
latency data we acquired from RIPE Atlas to correlate it with the intensity of
solar magnetic storms. Solar magnetic storms are usually identified by the Kp
index \cite{Bartels1957}. It is a number between zero and nine, where nine
indicates the strongest kind of storm. We found data historic Kp indexes on the
website of the Helmholtz-Zentrum Potsdam Deutsches GeoForschungsZentrum, G. F.
Z. \cite{GFZ2023}.

To get an understanding of the influences of solar magnetic storms on Starlink
performance, we will look at the intensity of the solar magnetic storm in
comparison to the TLS latency. We expect that with increasing intensity of
solar magnetic storms (i.e., higher Kp index), the latency will also increase.
Therefore, we used the average Kp index from GFZ database over a single day and
correlated it with the mean TLS handshake latency over each single day observed
in our dataset. The following correlation values resulted:

\begin{itemize}
	\item Pearson Correlation: $\approx 0.03$
	\item Kendall Correlation: $\approx 0.01$
	\item Spearman Correlation: $\approx 0.01$
\end{itemize}

One can see that the values are nearly zero. That implies that the dimensions
are nearly orthogonal. Therefore, we were not able to observe a correlation
between latency and Kp index. In its extreme, the data does not correlate at
all and observations form Ma~et~al.~\cite{DBLP:conf/infocom/MaCZCML23} are most
likely incorrect.

\begin{takeaway}{Correlation of Starlink Latency and Solar Magnetic Storms}
	Starlink TLS handshake latency and solar magnetic storms do not
	correlate.
\end{takeaway}

